---
title: "STAT 331 Portfolio"
author: "Aidan Lui"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute:
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an A.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1

```{r}
#| label: wd-1-csv-1
# Lab 3

data <- read.csv(here("week2", "lab3", "teacher_evals.csv"))
```

-   `csv` Example 2

```{r}
#| label: wd-1-csv-2
# Lab 2

surveys <- read_csv(here("week1", "lab2", "surveys.csv"))
```

-   `xlsx`

```{r}
#| label: wd-1-xlsx
# PA 4

military <- read_xlsx("gov_spending_per_capita.xlsx",
                      sheet = "Share of Govt. spending",
                      skip  = 7,
                      n_max = 190,
                      na = c("..", "xxx", ". .")
                      )
```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns

```{r}
#| label: wd-2-ex-1
# Lab 4 Question 3

tax_rev_join <- tax_rev |>
  select(entity_name, year, total_property_taxes, sales_and_use_taxes)

ca_childcare <- ca_childcare |>
  left_join(
    tax_rev_join,
    by = c("county_name" = "entity_name", "study_year" = "year")
  )

ca_childcare
```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2
# Lab 3 Question 5 Revised

teacher_evals_clean <- data |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    across(c(teacher_id, course_id), as.character),
    across(c(sex, academic_degree), as.factor)
  ) |>
    select(-c(
    stud_grade_avg, stud_grade_std, stud_grade_var_coef,
    percent_failed, stud_grade_avg_cur, stud_grade_std_cur,
    stud_grade_var_coef_cur, class_duration, weekday, time_of_day,
    SET_score_1sem, maximum_score, gender))
```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`)

```{r}
#| label: wd-2-ex-3
# Challenge 4

ca_table <- ca_summary |>
  summarize(
    across(
      starts_with(c("mc_", "mfcc_")), 
      median,
      na.rm = TRUE
))

ca_table <- tibble(
  Age_Group = c("Infant", "Toddler", "Preschool"),
  Center_Based_Median = scales::dollar(c(ca_table$mc_infant, ca_table$mc_toddler, ca_table$mc_preschool)),
  Family_In_Home_Median = scales::dollar(c(ca_table$mfcc_infant, ca_table$mfcc_toddler,                                                                  ca_table$mfcc_preschool)),
  Difference = scales::dollar(c(
                              ca_table$mc_infant - ca_table$mfcc_infant,
                              ca_table$mc_toddler - ca_table$mfcc_toddler,
                              ca_table$mc_preschool - ca_table$mfcc_preschool)
  )
)

ca_table
```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-3-numeric-ex-1
# Lab 3 Question 10

#Growing Comment: You should only call the filter() function once!

teacher_evals_clean |>
  filter(question_no == 901) |>
  group_by(teacher_id) |>
  summarize(
    avg_q1 = mean(SET_score_avg, na.rm = TRUE),
    n_courses = n_distinct(course_id),
    .groups = "drop") |>
  filter(n_courses >= 5) |>
  filter(avg_q1 == max(avg_q1, na.rm = TRUE) |
         avg_q1 == min(avg_q1, na.rm = TRUE)) |>
  arrange(desc(avg_q1))
```

-   Numeric Example 2

```{r}
#| label: wd-3-numeric-ex-2
# Lab 4 Question 5

median_income_by_region <- ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(region, study_year) |>
  summarize(
    value = median(mhi_2018, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = study_year,
    values_from = value
  ) |>
  arrange(desc(`2018`))

median_income_by_region
```

-   Character Example 1 (any context)

```{r}
#| label: wd-3-character
# Lab 5 

w1 <- person |>
  filter(str_detect(address_street_name,"Northwestern Dr")) |>
  filter(address_number == max(address_number))

w2 <- person |>
  filter(str_detect(address_street_name,"Franklin Ave"),
        str_detect(name,("^Annabel"))
  )

w1$name
w2$name
```

-   Character Example 2 (example must use functions from **stringr**)

```{r}
#| label: wd-3-string
# Lab 5

w1 <- person |>
  filter(str_detect(address_street_name,"Northwestern Dr")) |>
  filter(address_number == max(address_number))

w2 <- person |>
  filter(str_detect(address_street_name,"Franklin Ave"),
        str_detect(name,("^Annabel"))
  )

w1$name
w2$name
```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date
# Lab 5

gfn <- get_fit_now_member |>
  filter(str_detect(id, "^48Z"),
         str_detect(membership_status,"gold"))

gfn_checkins <- get_fit_now_check_in |>
  filter(membership_id %in% gfn$id) |>
  mutate(date = ymd(as.character(check_in_date)))|>
  filter(date == ymd("20180109"))

gfn_people <- gfn_checkins |>
  inner_join(gfn, by = c("membership_id" = "id")) |>
  inner_join(person, by = c("person_id" = "id")) 

gfn_suspect <- gfn_people |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(str_detect(plate_number, "H42W"))

gfn_suspect$name.x
```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-4-numeric-ex-1
# Challenge 3 Question 1

teacher_evals_compare <- data |>
  filter(question_no == 903) |>
  mutate(
    set_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
    sen_level = case_when(
      seniority <= 4 ~ "junior",
      seniority >= 5 & seniority <= 8 ~ "senior",
      seniority > 8 ~ "very senior"
    )
  ) |>
  select(course_id, set_level, sen_level)
```

-   Numeric Example 2

```{r}
#| label: wd-4-numeric-ex-2
# Challenge 7 Question 1
#Growing Comment: Why not name the column lenght_cm? It seems confusing to name it just length when the other column has units. Also, do you need to use mutate when you are just renaming a column?

fish <- fish |>
  rename(length_mm = length) |>
  mutate(length_cm = length_mm/10)
```

-   Factor Example 1 (renaming levels)

```{r}
#| label: wd-4-factor-ex-1
#Challenge 4

plot_data <- ca_summary |>
  select(mc_infant, mc_toddler, mc_preschool,
         mfcc_infant, mfcc_toddler, mfcc_preschool) |>
  pivot_longer(
    cols = everything(),
    names_to = "age_group",
    values_to = "weekly_price"
  ) |>
  mutate(
    age_group = recode(
      age_group,
      mc_infant    = "Center_Infant",
      mc_toddler   = "Center_Toddler",
      mc_preschool = "Center_Preschool",
      mfcc_infant     = "Family_Infant",
      mfcc_toddler    = "Family_Toddler",
      mfcc_preschool  = "Family_Preschool"
    )
  )
```

-   Factor Example 2 (reordering levels)

```{r}
#| label: wd-4-factor-ex-2
#Challenge 4

plot_data <- ca_summary |>
  select(mc_infant, mc_toddler, mc_preschool,
         mfcc_infant, mfcc_toddler, mfcc_preschool) |>
  pivot_longer(
    cols = everything(),
    names_to = "age_group",
    values_to = "weekly_price"
  ) |>
  mutate(
    age_group = recode(
      age_group,
      mc_infant = "Center_Infant",
      mc_toddler = "Center_Toddler",
      mc_preschool = "Center_Preschool",
      mfcc_infant = "Family_Infant",
      mfcc_toddler = "Family_Toddler",
      mfcc_preschool = "Family_Preschool"
    ),
    setting = if_else(str_detect(age_group, "Center"), "Center-Based", "Family (In-Home)"),
    age_group = case_when(
      str_detect(age_group, "Infant") ~ "Infant",
      str_detect(age_group, "Toddler") ~ "Toddler",
      str_detect(age_group, "Preschool") ~ "Preschool"
    ),
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool")
  )
```

-   Character (example must use functions from **stringr**)

```{r}
#| label: wd-4-string
#Lab 4 Question 7

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price)
  )
```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date
# Lab 5 Revised

concert_people <- facebook_event_checkin |>
  mutate(date = ymd(date)) |>
  filter(
    event_name == "SQL Symphony Concert",
    year(date) == 2017,
    month(date) == 12
  ) |>
  count(person_id) |>
  filter(n == 3)

mastermind <- person |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(
    str_detect(hair_color, "red"),
    height >= 65 & height <= 67,
    str_detect(car_make, "Tesla"),
    str_detect(car_model, "Model S")
  ) |>
  inner_join(concert_people, by = c("id" = "person_id"))

mastermind$name
```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1

```{r}
#| label: wd-5-left-ex-1
# Lab 4 Question 3

tax_rev_join <- tax_rev |>
  select(entity_name, year, total_property_taxes, sales_and_use_taxes)

ca_childcare <- ca_childcare |>
  left_join(
    tax_rev_join,
    by = c("county_name" = "entity_name", "study_year" = "year")
  )

ca_childcare
```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right
# Lab 5 Revised

witness_info <- person |>
  filter(
    (address_street_name == "Northwestern Dr" &
     address_number == max(address_number)) |
    (str_detect(name, "Annabel") &
     address_street_name == "Franklin Ave")
  ) |>
  right_join(interview, by = c("id" = "person_id")) |>
  filter(!is.na(name))

witness_info$name
```

-   `left_join()` **or** `right_join()` Example 2

```{r}
#| label: wd-5-left-right-ex-2
# Challenge 4

ca_summary <- childcare_costs |>
  left_join(counties, by = "county_fips_code") |>
  filter(state_name == "California") 
```

-   `inner_join()` Example 1

```{r}
#| label: wd-5-inner-ex-1
# Lab 5

gfn <- get_fit_now_member |>
  filter(str_detect(id, "^48Z"),
         str_detect(membership_status,"gold"))

gfn_checkins <- get_fit_now_check_in |>
  filter(membership_id %in% gfn$id) |>
  mutate(date = ymd(as.character(check_in_date)))|>
  filter(date == ymd("20180109"))

gfn_people <- gfn_checkins |>
  inner_join(gfn, by = c("membership_id" = "id")) |>
  inner_join(person, by = c("person_id" = "id")) 

gfn_suspect <- gfn_people |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(str_detect(plate_number, "H42W"))

gfn_suspect$name.x
```

-   `inner_join()` Example 2

```{r}
#| label: wd-5-inner-ex-2
# Lab 4 Question 2 

ca_childcare <- childcare_costs |>
  inner_join(counties, by = "county_fips_code") |>
  filter(state_name == "California") |>
  arrange(county_fips_code, study_year)

ca_childcare
```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi
# Lab 5 Revised

witness_interviews <- interview |>
  semi_join(witness_info, by = c("person_id" = "id"))

witness_interviews$transcript
```

-   `anti_join()`

```{r}
#| label: wd-6-anti
# Lab 5 Revised

gfn_exclude_members <- get_fit_now_member |>
  filter(!str_detect(id, "^48Z") | !str_detect(membership_status, "gold")) |>
  select(id)

gfn_checkins <- get_fit_now_check_in |>
  anti_join(gfn_exclude_members, by = c("membership_id" = "id")) |>
  mutate(date = ymd(as.character(check_in_date))) |>
  filter(date == ymd("20180109"))

gfn_people <- gfn_checkins |>
  inner_join(gfn, by = c("membership_id" = "id")) |>
  inner_join(person, by = c("person_id" = "id")) 

gfn_suspect <- gfn_people |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(str_detect(plate_number, "H42W"))

gfn_suspect$name
```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long
# Lab 4 Question 7

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price))
```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide
# Lab 4 Question 5 Revised

median_income_by_region <- ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(region, study_year) |>
  summarize(median_income = median(mhi_2018, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(
    names_from  = study_year,
    values_from = median_income,
    names_glue  = "Median_Income_{study_year}"
  )

median_income_by_region
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Lab 2
-   Lab 3
-   Lab 7 (revised: used the here() package)
-   Challenge 7 (revised: used the here() package)
-   Challenge 3 (revised: used the here() package)

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1
# Challenge 4

#| label: creating-a-plot
#| fig-width: 11
#| fig-height: 5

plot_data <- ca_summary |>
  select(mc_infant, mc_toddler, mc_preschool,
         mfcc_infant, mfcc_toddler, mfcc_preschool) |>
  pivot_longer(
    cols = everything(),
    names_to = "age_group",
    values_to = "weekly_price"
  ) |>
  mutate(
    age_group = recode(
      age_group,
      mc_infant = "Center_Infant",
      mc_toddler = "Center_Toddler",
      mc_preschool = "Center_Preschool",
      mfcc_infant = "Family_Infant",
      mfcc_toddler = "Family_Toddler",
      mfcc_preschool = "Family_Preschool"
    ),
    setting = if_else(str_detect(age_group, "Center"), "Center-Based", "Family (In-Home)"),
    age_group = case_when(
      str_detect(age_group, "Infant") ~ "Infant",
      str_detect(age_group, "Toddler") ~ "Toddler",
      str_detect(age_group, "Preschool") ~ "Preschool"
    ),
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool")
  )

custom_colors <- colorRampPalette(brewer.pal(8, "Accent"))(10)

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2

plot_data <- ca_summary |>
  select(mc_infant, mc_toddler, mc_preschool,
         mfcc_infant, mfcc_toddler, mfcc_preschool) |>
  pivot_longer(
    cols = everything(),
    names_to = "age_group",
    values_to = "weekly_price"
  ) |>
  mutate(
    age_group = recode(
      age_group,
      mc_infant = "Center_Infant",
      mc_toddler = "Center_Toddler",
      mc_preschool = "Center_Preschool",
      mfcc_infant = "Family_Infant",
      mfcc_toddler = "Family_Toddler",
      mfcc_preschool = "Family_Preschool"
    ),
    setting = if_else(str_detect(age_group, "Center"), "Center-Based", "Family (In-Home)"),
    age_group = case_when(
      str_detect(age_group, "Infant") ~ "Infant",
      str_detect(age_group, "Toddler") ~ "Toddler",
      str_detect(age_group, "Preschool") ~ "Preschool"
    ),
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool")
  )
```

-   Example of function formatting

```{r}
#| label: r-2-3
# Lab 8 Question 7

style_plot <- function(plot, theme_fn = theme_minimal, 
                       palette = brewer.pal(8, "Accent"), 
                       aesthetic = "color",
                       continuous = FALSE) {
  
    if (aesthetic == "fill") {
    if (continuous) {
      styled <- plot +
        scale_fill_gradient(low = palette[1], high = palette[2]) +
        theme_fn()
    } else {
      styled <- plot +
        scale_fill_manual(values = palette) +
        theme_fn()
    }
    
  } else if (aesthetic == "color") {
    if (continuous) {
      styled <- plot +
        scale_color_gradient(low = palette[1], high = palette[2]) +
        theme_fn()
    } else {
      styled <- plot +
        scale_color_manual(values = palette) +
        theme_fn()
    }
    
  } else {
    stop('aesthetic must be either "fill" or "color".')
  }
  
  return(styled)
}
```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example
# Lab 8 Question 3

plot_xy <- function(df, x, y) {
  x_name <- as_label(ensym(x))
  y_name <- as_label(ensym(y))
  
  x_label <- x_name |>
    str_replace_all("_", " ") |>
    str_to_title() |>
    str_replace(" G$", " (g)") |>
    str_replace(" Mm$", " (mm)")
  
  y_label <- y_name |>
    str_replace_all("_", " ") |>
    str_to_title() |>
    str_replace(" G$", " (g)") |>
    str_replace(" Mm$", " (mm)")
  
  ggplot(df, aes(x = {{ x }}, y = {{ y }}, color = {{ x }})) +
    geom_point() +
    labs(
      title = paste("Relationship between", y_label, "and", x_label),
      x = x_label,
      y = y_label,
      color = x_label
    ) +
    theme_bw() +
    theme(legend.position = "none") 
}
```

-   Example (function stops)

```{r}
#| label: r-3-function-stops
# Challenge 7 Question 4
#Growing Comment: Where are your function checks? What types of inputs does your function expect?

condition <- function(weight_g, length_cm) {
  if (!is.numeric(weight_g)) {
    stop("weight_g must be numeric")
  }
  if (!is.numeric(length_cm)) {
    stop("length_cm must be numeric")
  }
  if (any(length_cm == 0, na.rm = TRUE)) {
    stop("length_cm cannot contain zero")
  }
  
  (weight_g / (length_cm^3)) * 100
}
```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num
# Lab 10 Question 13

set.seed(1229)

ci_df |>
  slice_sample(n = 100) |>
  mutate(sim_id = row_number()) |>
  ggplot(aes(y = sim_id)) +
  geom_errorbar(aes(xmin = ci_low, xmax = ci_high, color = factor(cover)),
                 width = 2, linewidth = 0.8) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "black", linewidth = 0.2) +
  scale_color_manual(values = c("0" = "red", "1" = "skyblue3")) +
  labs(
    title = "95% Confidence Intervals of Estimated Regression Slope",
    subtitle = "CI's that do not include the population slope are highlighted <b style='color:red;'>red</b>.",
    x = "Slope Coefficient Estimate",
    y = ""
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
    plot.subtitle = element_markdown(hjust = 0.5, size = 12),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
ci_df
```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat
# Lab 2 Question 16

ggplot(data = surveys,
       mapping = aes(x = weight, y = species)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(
    color = "darkred",
    alpha = 0.5,
    width = 0.2
  ) +
    labs(
      x = "Weight of Individuals (grams)",
      y = "Species",
      title = "Distribution of Individual Weights Across Species"
    ) +
    theme_bw()
```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat
# Challenge 3 Question 2

teacher_evals_compare |>
  ggplot(aes(x = sen_level, fill = set_level)) +
  geom_bar(position = "fill", width = 0.9) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.25),
    labels = c("0%", "25%", "50%", "75%", "100%")
  ) +
  scale_fill_manual(
    values = c("excellent"="#B69AD6", "standard"="#A97B44"),
    name   = "Evaluation Rating"
  ) +
  labs(
    title = "Evaluation of Teachers' Use of Activities",
    x = "Years of Experience",
    y = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title   = element_text(size = 22, face = "plain", hjust = 0),
    legend.position = "top",
    legend.title = element_text(size = 16),
    legend.text  = element_text(size = 14),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
  )
```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date
# Lab 4 Question 6

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price)
  ) |>
  ggplot(aes(study_year, median_weekly_price, color = region_label)) +
  geom_point(alpha = 0.6) +
  geom_smooth(span = 0.9) +
  facet_wrap(~ age_group, ncol = 3) +
  scale_x_continuous(breaks = seq(2008, 2018, 2), limits = c(2008, 2018)) +
  scale_y_continuous(breaks = seq(100, 500, 100), limits = c(100, 500)) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare",
    x = "Study Year",
    y = "Weekly Median Price",
    color = "California"
  ) +
  theme_minimal() +
  theme(
    aspect.ratio = 1,
    legend.position = "right",
    panel.border = element_rect(color = "black", fill = NA, size = 0.6),
    strip.background = element_rect(
      fill = "grey85", color = "grey50", size = 0.6)
  )
```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1
# Lab 4 Question 7

#| fig-width: 11
#| fig-height: 6

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price)
  ) |>
  ggplot(aes(study_year, median_weekly_price, color = region_label)) +
  geom_point(alpha = 0.6) +
  geom_smooth(span = 0.9) +
  facet_wrap(~ age_group, ncol = 3) +
  scale_x_continuous(breaks = seq(2008, 2018, 2), limits = c(2008, 2018)) +
  scale_y_continuous(breaks = seq(100, 500, 100), limits = c(100, 500)) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare",
    x = "Study Year",
    y = "Weekly Median Price",
    color = "California"
  ) +
  theme_minimal() +
  theme(
    aspect.ratio = 1,
    legend.position = "right",
    panel.border = element_rect(color = "black", fill = NA, size = 0.6),
    strip.background = element_rect(
      fill = "grey85", color = "grey50", size = 0.6)
  )
```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2
# Challenge 2 Hot Revised

ggplot(data = surveys,
       mapping = aes(x = weight, y = species, color = sex)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha = 0.5, width = 0, height = 0.15) +
  labs(
    x = "Weight of Individuals (grams)",
    y = "Species",
    title = "Distribution of Weights Across Species of Rodents by Sex",
    subtitle = "Colored by <span style='color:#0072B2;'>Male</span> and <span   style='color:#E69F00;'>Female</span>"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(hjust = 1),
    legend.position = "none",
    plot.subtitle = element_markdown()
  ) +
  scale_color_manual(values = c("M" = "#0072B2", "F" = "#E69F00"))
```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3
# Challenge 4

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4
# Challenge 4

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5
# Lab 4 Question 7

#| fig-width: 11
#| fig-height: 6

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price)
  ) |>
  ggplot(aes(study_year, median_weekly_price, color = region_label)) +
  geom_point(alpha = 0.6) +
  geom_smooth(span = 0.9) +
  facet_wrap(~ age_group, ncol = 3) +
  scale_x_continuous(breaks = seq(2008, 2018, 2), limits = c(2008, 2018)) +
  scale_y_continuous(breaks = seq(100, 500, 100), limits = c(100, 500)) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare",
    x = "Study Year",
    y = "Weekly Median Price",
    color = "California"
  ) +
  theme_minimal() +
  theme(
    aspect.ratio = 1,
    legend.position = "right",
    panel.border = element_rect(color = "black", fill = NA, size = 0.6),
    strip.background = element_rect(
      fill = "grey85", color = "grey50", size = 0.6)
  )
```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1
#Challenge 4

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2
#Challenge 2 Medium Revised

#Growing Comment: I can't see the boxplots at all, only the jittered points. I'd like you to play around with ways to have both layers in the plot but for the reader to see both of them. This can be changing the ordering of the layers or increasing your transparency. The jittered points need to be separated by the boxplot. I'd recommend playing with the jittering so the points look like they belong to the correct boxplot! 

colors <- colorRampPalette(brewer.pal(8, "Accent"))(10)

ggplot(data = surveys, mapping = aes(x = weight, y = species, color = sex)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8) +
  geom_jitter(
    aes(color = sex),
    width = 0.2,
    alpha = 0.5,
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("M" = colors[5], "F" = colors[8])) +
  labs(
    x = "Body Mass (grams)",
    y = "Species",
    title = "Weight Distribution Across Species of Rodents"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 13, hjust = 0.5),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", size = 1)
  )
```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2
# Challenge 4

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3
# Challenge 4

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize
# Lab 3 Question 7

instructor_demo <- teacher_evals_clean |>
  distinct(teacher_id, .keep_all = TRUE) |>
  select(teacher_id, academic_degree, seniority, sex)

instructor_demo |>
  count(academic_degree)

instructor_demo |>
  summarize(
    avg_seniority = mean(seniority, na.rm = TRUE),
    min_seniority = min(seniority, na.rm = TRUE),
    max_seniority = max(seniority, na.rm = TRUE)
  )

instructor_demo |>
  count(sex)
```

-   Example using `across()`

```{r}
#| label: dvs-4-across
# Lab 7 Question 1
#Growing Comment: It looks like you are using colSums(). This is a tool that was not explicitly taught in lecture / lab. While this tool accomplishes this task, that is not the intention of this assignment. You are being assessed on your understanding of the tools that have been taught in this course. In the future, please use tools that have been taught in the course when addressing Lab questions.

fish_missing <- fish |>
  summarize(across(everything(), ~sum(is.na(.x)), .names = "{.col}_na"))

fish_missing
```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1
# Lab 4 Question 5

median_income_by_region <- ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(region, study_year) |>
  summarise(
    value = median(mhi_2018, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = study_year,
    values_from = value
  ) |>
  arrange(desc(`2018`))

median_income_by_region
```

-   Example 2

```{r}
#| label: dvs-5-2
# Lab Question 12

teacher_evals_clean |>
  filter(academic_degree %in% c("dr", "prof")) |>
  group_by(teacher_id, seniority, sex) |>
  summarize(
    avg_resp = mean(resp_share, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(avg_resp == max(avg_resp, na.rm = TRUE) |
         avg_resp == min(avg_resp, na.rm = TRUE)) |>
  arrange(desc(avg_resp))
```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1
# Challenge 4

ca_table <- ca_summary |>
  summarize(
    across(
      starts_with(c("mc_", "mfcc_")), 
      median,
      na.rm = TRUE
))

ca_table <- tibble(
  Age_Group = c("Infant", "Toddler", "Preschool"),
  Center_Based_Median = scales::dollar(c(ca_table$mc_infant, ca_table$mc_toddler, ca_table$mc_preschool)),
  Family_In_Home_Median = scales::dollar(c(ca_table$mfcc_infant, ca_table$mfcc_toddler,                                                                  ca_table$mfcc_preschool)),
  Difference = scales::dollar(c(
                              ca_table$mc_infant - ca_table$mfcc_infant,
                              ca_table$mc_toddler - ca_table$mfcc_toddler,
                              ca_table$mc_preschool - ca_table$mfcc_preschool)
  )
)

ca_table
```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2
# Lab 9 Question 8

fish |>
  map_int(~ sum(is.na(.x))) |>
  enframe(name = "Variable", value = "Missing") |>
  arrange(desc(Missing)) |>
  gt() |>
  tab_header(
    title = md("**Number of Missing Values for Fish Measurement**"),
    subtitle = "Data Collected from 1989 to 2006 on the Blackfoot Fish River in Montana"
  ) |>
  cols_label(
    Variable = md("**Measurment Variable**"),
    Missing  = md("**Missing Values**")
  ) |>
  fmt_number(columns = Missing, decimals = 0) |>

  tab_style(
    style = list(
      cell_fill(color = "#b6e3b0"),     
      cell_text(color = "black", weight = "bold")
    ),
    locations = cells_body(columns = Missing, rows = Missing == 0)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#f4a8a8"),         
      cell_text(color = "black", weight = "bold")
    ),
    locations = cells_body(columns = Missing, rows = Missing > 0)
  ) |>
  tab_options(
    table.font.color = "black",            
    table.width = pct(60)
  )
```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3
# Lab 3 Question 10

#Growing Comment: You should only call the filter() function once!

teacher_evals_clean |>
  filter(question_no == 901) |>
  group_by(teacher_id) |>
  summarize(
    avg_q1 = mean(SET_score_avg, na.rm = TRUE),
    n_courses = n_distinct(course_id),
    .groups = "drop") |>
  filter(n_courses >= 5) |>
  filter(avg_q1 == max(avg_q1, na.rm = TRUE) |
         avg_q1 == min(avg_q1, na.rm = TRUE)) |>
  arrange(desc(avg_q1))
```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1
# Lab 9 Question 8

fish |>
  map_int(~ sum(is.na(.x))) |>
  enframe(name = "Variable", value = "Missing") |>
  arrange(desc(Missing)) |>
  gt() |>
  tab_header(
    title = md("**Number of Missing Values for Fish Measurement**"),
    subtitle = "Data Collected from 1989 to 2006 on the Blackfoot Fish River in Montana"
  ) |>
  cols_label(
    Variable = md("**Measurment Variable**"),
    Missing  = md("**Missing Values**")
  ) |>
  fmt_number(columns = Missing, decimals = 0) |>

  tab_style(
    style = list(
      cell_fill(color = "#b6e3b0"),     
      cell_text(color = "black", weight = "bold")
    ),
    locations = cells_body(columns = Missing, rows = Missing == 0)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#f4a8a8"),         
      cell_text(color = "black", weight = "bold")
    ),
    locations = cells_body(columns = Missing, rows = Missing > 0)
  ) |>
  tab_options(
    table.font.color = "black",            
    table.width = pct(60)
  )
```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`)

```{r}
#| label: dvs-7-ex-2
# Lab 4 Question 5

median_income_by_region <- ca_childcare |>
  filter(study_year %in% c(2008, 2018)) |>
  group_by(region, study_year) |>
  summarise(
    value = median(mhi_2018, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = study_year,
    values_from = value
  ) |>
  arrange(desc(`2018`))

median_income_by_region

# Lab 7 Question 2
fish_table <- fish_missing |>
  rename_with(~ gsub("_na$", "", .x)) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Number of Missing Values"
  )

fish_table
```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call
# Lab 4 Question 7

ca_childcare |>
  select(region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  rename(
    Infant = mc_infant,
    Toddler = mc_toddler,
    Preschool = mc_preschool
  ) |>
  pivot_longer(
    c(Infant, Toddler, Preschool),
    names_to = "age_group",
    values_to = "median_weekly_price"
  ) |>
  mutate(
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool"),
    region_label = str_remove(region, "Region"),
    region_label = fct_reorder2(region_label, study_year, median_weekly_price))
```

-   using `across()`

```{r}
#| label: pe-1-across
# Lab 3 Question 5

teacher_evals_clean <- data |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(
    across(c(teacher_id, course_id), as.character),
    across(c(sex, academic_degree), as.factor)
  ) |>
  select(course_id, teacher_id, question_no, no_participants, resp_share, SET_score_avg, percent_failed_cur, academic_degree, seniority, sex)
```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1
# Lab 9 Question 4

evals <- evals |>
  map_if(~ is.character(.x), as.factor) |>
  bind_cols()
```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1
# Lab 8 Question 1

rescale_01 <- function(x) {
  if (!is.numeric(x)) {
    stop("Input vector must be numeric.")
  }

  if (length(x) <= 1) {
    stop("Input vector must have more than one value.")
  }

  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2
# Lab 8 Question 1

rescale_column <- function(data, cols) {
  data |>
    mutate(
      across(.cols = {{ cols }},
             .fns = rescale_01
            )
      )
}
```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3
# Lab 8 Question 3

plot_xy <- function(df, x, y) {
  x_name <- as_label(ensym(x))
  y_name <- as_label(ensym(y))
  
  x_label <- x_name |>
    str_replace_all("_", " ") |>
    str_to_title() |>
    str_replace(" G$", " (g)") |>
    str_replace(" Mm$", " (mm)")
  
  y_label <- y_name |>
    str_replace_all("_", " ") |>
    str_to_title() |>
    str_replace(" G$", " (g)") |>
    str_replace(" Mm$", " (mm)")
  
  ggplot(df, aes(x = {{ x }}, y = {{ y }}, color = {{ x }})) +
    geom_point() +
    labs(
      title = paste("Relationship between", y_label, "and", x_label),
      x = x_label,
      y = y_label,
      color = x_label
    ) +
    theme_bw() +
    theme(legend.position = "none") 
}
```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across
# Challenge 4

ca_table <- ca_summary |>
  summarize(
    across(
      starts_with(c("mc_", "mfcc_")), 
      median,
      na.rm = TRUE
))
```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1
# Lab 9 Question 2

surveys_types <- map_chr(surveys, ~ class(.x)) |> 
  enframe(name = "Variable", value = "Data Type")

kable(surveys_types, caption = "Data Types of Variables in the Surveys Dataset")
```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2
# Lab 9 Question 7

fish |>
  map_int(~ sum(is.na(.x))) |>
  enframe(name = "Variable", value = "Missing") |>
  arrange(desc(Missing))
```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1
# Lab 5

w1 <- person |>
  filter(str_detect(address_street_name,"Northwestern Dr")) |>
  filter(address_number == max(address_number))

w2 <- person |>
  filter(str_detect(address_street_name,"Franklin Ave"),
        str_detect(name,("^Annabel"))
  )

w1$name
w2$name
```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2
#Challenge 4

#| fig-width: 11
#| fig-height: 5

plot_data <- ca_summary |>
  select(mc_infant, mc_toddler, mc_preschool,
         mfcc_infant, mfcc_toddler, mfcc_preschool) |>
  pivot_longer(
    cols = everything(),
    names_to = "age_group",
    values_to = "weekly_price"
  ) |>
  mutate(
    age_group = recode(
      age_group,
      mc_infant = "Center_Infant",
      mc_toddler = "Center_Toddler",
      mc_preschool = "Center_Preschool",
      mfcc_infant = "Family_Infant",
      mfcc_toddler = "Family_Toddler",
      mfcc_preschool = "Family_Preschool"
    ),
    setting = if_else(str_detect(age_group, "Center"), "Center-Based", "Family (In-Home)"),
    age_group = case_when(
      str_detect(age_group, "Infant") ~ "Infant",
      str_detect(age_group, "Toddler") ~ "Toddler",
      str_detect(age_group, "Preschool") ~ "Preschool"
    ),
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool")
  )

custom_colors <- colorRampPalette(brewer.pal(8, "Accent"))(10)

ggplot(plot_data, aes(x = age_group, y = weekly_price, fill = age_group)) +
  geom_violin(trim = TRUE, alpha = 0.8, linewidth = 0.5) +
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "errorbar", width = 0.45, color = "black", linewidth = 0.5) +
  stat_summary(fun = median, geom = "text",
               aes(label = scales::dollar(after_stat(y))),
               vjust = -0.6, size = 3, color = "black") +
  facet_wrap(~ setting, ncol = 2) +
  scale_y_continuous(labels = scales::dollar) +
  scale_fill_manual(values = custom_colors) +   
  scale_color_manual(values = custom_colors) +
  labs(
    title = "Distribution of Full-Time Weekly Childcare Costs in California (2008–2018)",
    subtitle = "Violin shows distribution shape; bold line and label mark the median",
    x = "Age Group of Children",
    y = "Full-Time Weekly Cost"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.7),
    strip.background = element_rect(fill = "grey95", color = "grey70"),
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(face = "italic", color = "grey30"),
    axis.title = element_text(face = "bold"),
    axis.line = element_line(color = "grey60"),
    panel.grid.minor = element_blank()
  )
```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1
# Lab 10 Question 1 Revised

randomBabies <- function(n_babies) {
  random_order <- sample(1:n_babies, size = n_babies, replace = FALSE)
  sum(random_order == 1:n_babies)
}
```

-   Example 2

```{r}
#| label: dsm-1-2
# Lab 10 Question 8

mycifun <- function(beta0, beta1, n) {
  x  <- runif(n, min = 0, max = 1)
  ep <- rnorm(n, mean = 0, sd = 1)  
  y  <- beta0 + beta1 * x + ep
  data_obs <- data.frame(x = x, y = y)

  fit <- lm(y ~ x, data = data_obs)

  tt <- tidy(fit, conf.int = TRUE)
  slope_row <- tt[tt$term == "x", c("estimate", "conf.low", "conf.high")]

  cover <- as.integer(beta1 >= slope_row$conf.low & beta1 <= slope_row$conf.high)
}
```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1
# Lab 2 Question 17

species_mod <- aov(weight ~ species, data = surveys)
summary(species_mod)
```

-   Example 2

```{r}
#| label: dsm-2-2
# Challenge 3 Question 3

chisq.test(teacher_evals_compare$set_level, teacher_evals_compare$sen_level)
```

-   Example 3

```{r}
#| label: dsm-2-3
# Lab 4 Question 9

#Growing Comment: I strongly recommend against nested functions, as they are difficult for people to understand what your code is doing. Is this plot for only 2018?

reg_data <- ca_childcare |>
  filter(!is.na(mc_infant), !is.na(mhi_2018))

reg_mod1 <- lm(
  mc_infant ~ mhi_2018,
  data = reg_data
)

summary(reg_mod1)
```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

I revised my thinking by doing every revision for the labs and challenges and wrote a reflection for every chunk of code that I revised, no matter how small the fix was. Writing down what I changed helped the ideas stick because I wasn’t just fixing things, I was thinking about why they needed fixing. It also made me notice patterns in the mistakes I made, so I could avoid them in the future. Over time I started catching those issues before anyone even pointed them out. Doing this over and over made my code cleaner, and it helped me understand what good, consistent code should look like.

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

Throughout this course I have extended my thinking by not just understanding what was taught in class but going out of my way to look up potentially new or even better functions when the current functions I was using looked to repetitive or didn't an output to my liking. I paid attention to all the feedback that was giving to me through my graded assignments and code reviews. I always took into consideration and immediately went to fix my code while ideas were still fresh in my head. I even started to care not only about if the code ran but how it looked and read to other people especially small details like naimg convention for my variables.

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->

This is my feedback that I am proud of, from Lab 5: One thing that you did really well is keeping your code clear and easy to read. Each step builds nicely on the previous one, and you did a great job using line breaks after each pipe to make your workflow organized. The alignment of arguments inside functions like mutate() (for example, lining up is8, date_chr, and date_std) makes your logic easy to follow and shows strong attention to structure. The spacing between chunks also helps separate the major steps in your process, which makes it simple to see how the data is being transformed at each stage. Overall, the layout looks clean, deliberate, and professional.

One thing you can improve in code formatting is simplifying the syntax now that tidyverse automatically loads dplyr and lubridate. There’s no need to write dplyr::filter() or lubridate::year() since those functions are already available when using tidyverse. Removing the :: calls would make your pipelines cleaner and more consistent with tidyverse style. You can also make your filtering step more efficient by using the ymd() function instead of calling year(), month(), and day()

Doing peer reviews and paired programming helped me get better at communicating my ideas and talking through problems with other people. It made me realize that my way of thinking was sometimes too straightforward and inefficient. Hearing how other people approached problems gave me a new way of thinking of ideas. Listening to others allowed to come up with more creative ideas when I was working on my own.
